---
title: "Shiny Integration"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Shiny Integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

`gridify()` returns a grid graphical object, so it should be rendered in Shiny
using `plotOutput()` and `renderPlot()`.

`gridify_with_settings` is a compact module that wraps this pattern and adds
height/width sliders with PNG and PDF download buttons.

For more detail on layout options and `set_cell()` usage, see
`vignette("simple_examples", package = "gridify")`.

## Example with a `ggplot2` figure

Here we create a `ggplot2` figure and wrap it in a `gridify` object with
`pharma_layout_base()`.
We then render it in Shiny through `gridify_with_settings_srv()`.

```{r}
library(shiny)
library(ggplot2)
library(gridify)

ui <- fluidPage(
  titlePanel("gridify + Shiny"),
  gridify_with_settings_ui("demo")
)

server <- function(input, output, session) {
  gridify_r <- reactive({
    figure_obj <- ggplot2::ggplot(mtcars, aes(x = mpg, y = wt)) +
      ggplot2::geom_point()

    gridify(
      object = figure_obj,
      layout = pharma_layout_base()
    ) |>
      set_cell("header_left_1", "My Company") |>
      set_cell("header_left_2", "<PROJECT> / <INDICATION>") |>
      set_cell("header_left_3", "<STUDY>") |>
      set_cell("header_right_2", "<Draft or Final>") |>
      set_cell("output_num", "<Figure> xx.xx.xx") |>
      set_cell("title_1", "<Title 1>") |>
      set_cell("title_2", "<Title 2>") |>
      set_cell("note", "<Note or Footnotes>") |>
      set_cell("references", "<References:>") |>
      set_cell("footer_left", "Program: <PROGRAM NAME>, YYYY-MM-DD at HH:MM") |>
      set_cell("footer_right", "Page 1 of 1")
  })

  gridify_with_settings_srv("demo", gridify_r = gridify_r)
}

shinyApp(ui, server)
```

## Example with a base R figure

Here we create a base R plot as a formula object and pass it to `gridify()`.
We then apply `complex_layout()` and render it with
`gridify_with_settings_srv()`.

```{r}
library(shiny)
library(gridify)

ui <- fluidPage(
  titlePanel("gridify + Shiny + base R"),
  gridify_with_settings_ui("base_demo")
)

server <- function(input, output, session) {
  gridify_r <- reactive({
    formula_object <- ~ barplot(1:10)

    gridify(
      object = formula_object,
      layout = complex_layout()
    ) |>
      set_cell("header_left", "Left Header") |>
      set_cell("header_middle", "Middle Header") |>
      set_cell("header_right", "Right Header") |>
      set_cell("title", "Title") |>
      set_cell("subtitle", "Subtitle") |>
      set_cell("note", "Note") |>
      set_cell("footer_left", "Left Footer") |>
      set_cell("footer_middle", "Middle Footer") |>
      set_cell("footer_right", "Right Footer")
  })

  gridify_with_settings_srv("base_demo", gridify_r = gridify_r)
}

shinyApp(ui, server)
```

## Example with a `flextable`

Here we create a `flextable` object and pass it to `gridify()` with
`pharma_layout_letter()`.
We then render it in Shiny using `gridify_with_settings_srv()`.

```{r}
library(shiny)
library(flextable)
library(gridify)

ui <- fluidPage(
  titlePanel("gridify + Shiny + flextable"),
  gridify_with_settings_ui("ft_demo")
)

server <- function(input, output, session) {
  gridify_r <- reactive({
    ft <- flextable::flextable(head(mtcars[1:10]))

    gridify(
      object = ft,
      layout = pharma_layout_letter()
    ) |>
      set_cell("header_left_1", "My Company") |>
      set_cell("header_left_2", "<PROJECT> / <INDICATION>") |>
      set_cell("header_left_3", "<STUDY>") |>
      set_cell("header_right_2", "<Draft or Final>") |>
      set_cell("output_num", "<Table> xx.xx.xx") |>
      set_cell("title_1", "<Title 1>") |>
      set_cell("title_2", "<Title 2>") |>
      set_cell("note", "<Note or Footnotes>") |>
      set_cell("references", "<References:>") |>
      set_cell("footer_left", "Program: <PROGRAM NAME>, YYYY-MM-DD at HH:MM") |>
      set_cell("footer_right", "Page 1 of 1")
  })

  gridify_with_settings_srv("ft_demo", gridify_r = gridify_r)
}

shinyApp(ui, server)
```

## Example with a `gt` table

Here we create a `gt` table and pass it to `gridify()` with
`pharma_layout_letter()`.
The server then wraps this object with `gridify_with_settings_srv()` for
interactive rendering and download.

```{r}
library(shiny)
library(gt)
library(gridify)

ui <- fluidPage(
  titlePanel("gridify + Shiny + gt"),
  gridify_with_settings_ui("gt_demo")
)

server <- function(input, output, session) {
  gridify_r <- reactive({
    gt_tbl <- gt::gt(head(mtcars[1:10]))

    gridify(
      object = gt_tbl,
      layout = pharma_layout_letter()
    ) |>
      set_cell("header_left_1", "My Company") |>
      set_cell("header_left_2", "<PROJECT> / <INDICATION>") |>
      set_cell("header_left_3", "<STUDY>") |>
      set_cell("header_right_2", "<Draft or Final>") |>
      set_cell("output_num", "<Table> xx.xx.xx") |>
      set_cell("title_1", "<Title 1>") |>
      set_cell("title_2", "<Title 2>") |>
      set_cell("note", "<Note or Footnotes>") |>
      set_cell("references", "<References:>") |>
      set_cell("footer_left", "Program: <PROGRAM NAME>, YYYY-MM-DD at HH:MM") |>
      set_cell("footer_right", "Page 1 of 1")
  })

  gridify_with_settings_srv("gt_demo", gridify_r = gridify_r)
}

shinyApp(ui, server)
```

