---
title: "Transparency with gridify"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Transparency with gridify}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

## Introduction

It is assumed that you have already read the other vignettes and understand `gridify`.

The `gridify` package simplifies the creation of complex [grid](https://cran.r-project.org/package=grid)-based layouts in R. 
It builds on the base R `grid` package, providing a more intuitive interface for constructing advanced graphical arrangements—such as titles, headers, and footers—without requiring deep knowledge of `grid` itself. 
A good [introduction to `grid` is available here](https://bookdown.org/rdpeng/RProgDA/the-grid-package.html).

A key strength of `gridify` is its _meta-programming_ approach: when you print and assign a `gridify` object, you can retrieve and inspect the exact `grid` call code. 
This helps ensure **transparency** so you can see precisely how your outputs are composed. 
Such a feature is invaluable in regulated industries and any environment where you need to demonstrate how each output is generated.

## Example Usage

Below is a simple example showcasing how `gridify` wraps a `ggplot2` figure with headers, footers, and additional text elements, 
all while exposing the raw `grid` code used behind the scenes:

```{r, fig.width=7, fig.height=7}
library(gridify)
# For the native pipe (|>), R 4.1.0 or higher is recommended.
# Otherwise, you can use magrittr's %>%.
library(magrittr)

library(ggplot2)

# Create and print a gridify object
grid_call_behind <- gridify(
  object = ggplot2::ggplot(
    data = mtcars,
    ggplot2::aes(x = mpg, y = wt)
  ) +
    ggplot2::geom_line(),
  layout = complex_layout(scales = "fixed")
) %>%
  set_cell("header_left", "Left Header") %>%
  set_cell("header_middle", "Middle Header") %>%
  set_cell("header_right", "Right Header") %>%
  set_cell("title", "Title") %>%
  set_cell("subtitle", "Subtitle") %>%
  set_cell("note", "Note") %>%
  set_cell("footer_left", "Left Footer") %>%
  set_cell("footer_middle", "Middle Footer") %>%
  set_cell("footer_right", "Right Footer") %>%
  print()

# The underlying grid call is returned invisibly by the print() method.
# We can inspect or store it for further manipulations:
grid_call_behind
```

```{r, fig.width=7, fig.height=7}
# Retrieve the main object- either a figure or table
OBJECT <- attr(grid_call_behind, "env")[["OBJECT"]]
# Redraw the graphics by eval of the grid call
grid::grid.draw(eval(grid_call_behind))
```

Both of the graphics are the same.

When you call `print()` on a `gridify` object and assign it to a variable, `gridify` constructs and returns the `grid` call. 
Through meta-programming, `gridify` captures all `grid` calls required to produce your graphics.

## Key Transparency Benefits

1. **Direct Visibility into the Layout**  
   By exposing the raw `grid` code, `gridify` ensures you can see exactly how your figures are constructed. 
   This clear trail of code is especially helpful for validating complex graphics or meeting regulatory requirements.

2. **Educational Insight**  
   If you’re interested in learning how `grid` works, you can study the generated code. 
   It’s an excellent, hands-on way to build your knowledge of low-level graphics in R.

3. **Audit and Debugging**  
   The ability to retrieve the full `grid` calls makes it straightforward to diagnose layout issues or confirm that all elements are placed correctly.

## Conclusion

By providing direct access to the underlying `grid` code, `gridify` prioritizes **transparency**. 

For more details on other aspects like reproducibility, consistency, and advanced layouts, check out our package documentation and additional vignettes.
